---
title: "Data Cleaning & Modeling"
author: "Ishara Wijayaratne"
format:
  html:
    toc: true
    toc-location: left
    toc-title: "Contents"
    toc-depth: 3
editor: visual
---

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
library(lubridate) # used in function mutate()
library(gt) # to get a tabular format for the summary statistics
library(tibble)

#scatter plot matrix
library(ggplot2)
library(GGally)
library(patchwork)

#modeling
library(lme4)
library(sandwich)
library(lmtest)
```

## ride_full (the complete data set)

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_full <- read_excel("Ride Requests_2026-06-09.xlsx")

ride_full <- ride_full |>
  select(
    Request_Creation_Time    = `Request Creation Time`,
    Request_Status           = `Request Status`, #should be Completed
    Rider_ID                 = `Rider ID`,
    Booking_Method           = `Booking Method`, #any booking method is accepted
    Origin_Address           = `Origin Address`,
    Origin_Lat               = `Origin Lat`,
    Origin_Lng               = `Origin Lng`,
    Destination_Address      = `Destination Address`,
    Destination_Lat          = `Destination Lat`,
    Destination_Lng          = `Destination Lng`,
    
    Ride_Distance                 = `Ride Distance`,
    Pickup_Walking_Distance_ft    = `Pickup Walk Distance`,
    Dropoff_Walking_Distance_ft   = `Dropoff Walk Distance`,
    
    Ride_Duration            = `Ride Duration`, #amount of time the ride lasted
    Waiting_time_to_pickup_minutes = `On Demand Pickup ETA Minutes`, #Number of minutes from ride request to the estimated time of arrival (ETA)
    Deviation_minutes        = `On Demand Pickup ETA Deviation Minutes` #captures pickup delay, +ve a delay, -ve early
  )

ride_full <- ride_full |>
  mutate(across(c(Ride_Distance,
                  Pickup_Walking_Distance_ft,
                  Dropoff_Walking_Distance_ft,
                  Ride_Duration,
                  Waiting_time_to_pickup_minutes,
                  Deviation_minutes),
                as.numeric))


ride_full <- as.data.frame(ride_full)

ride_full <- ride_full |>
  mutate(Rider_ID = str_extract(as.character(Rider_ID), "\\d{7}"))

ride_full <- ride_full |>
  mutate(
    Request_Creation_Time = force_tz(as_datetime(Request_Creation_Time), "America/New_York"),
    creation_date         = as_date(Request_Creation_Time),
    creation_time         = hms::as_hms(Request_Creation_Time),
    study_week            = case_when(
      creation_date >= ymd("2026-03-30") & creation_date <= ymd("2026-04-04") ~ "Mar30-Apr4",
      creation_date >= ymd("2026-04-06") & creation_date <= ymd("2026-04-11") ~ "Apr 6-11",
      creation_date >= ymd("2026-04-13") & creation_date <= ymd("2026-04-18") ~ "Apr 13-18",
      creation_date >= ymd("2026-04-20") & creation_date <= ymd("2026-04-25") ~ "Apr 20-25",
      creation_date >= ymd("2026-04-27") & creation_date <= ymd("2026-05-02") ~ "Apr27-May2",
      creation_date >= ymd("2026-05-04") & creation_date <= ymd("2026-05-09") ~ "May 4-9",
      creation_date >= ymd("2026-05-11") & creation_date <= ymd("2026-05-16") ~ "May 11-16",
      creation_date >= ymd("2026-05-18") & creation_date <= ymd("2026-05-23") ~ "May 18-23",
      creation_date >= ymd("2026-05-25") & creation_date <= ymd("2026-05-30") ~ "May 25-30",
      TRUE ~ NA_character_
    )
  ) |>
  arrange(Rider_ID, creation_date, creation_time) |>  # order to the second
  group_by(Rider_ID, creation_date) |>
  mutate(first_ride_of_day = row_number() == 1L) |>
  ungroup() |>
  select(-Request_Creation_Time)   

ride_full <- ride_full |>
  relocate(creation_date, creation_time, Ride_Duration, Rider_ID)

glimpse(ride_full)
```

## ride_data (survey/qualtrics data set)

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

raw <- read_csv(
  "USE__May 25 2026__Wilson_SLJLMS .csv",
  skip = 3,                 # drop 3 top rows
  col_names = c(            # variable names
    "start_date_time", "end_date", "IPaddress", "progress", "duration",
    "finished", "recorded_date", "response_ID", "recipient_last_name",
    "latitude", "longitude", "Q_recaptcha_score", "phone_NO",
    "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday",
    "trip_purpose", "trip_purpose_text", "comments",
    "Q_recaptcha_status", "Q_recaptcha_error", "day", "via", "incentive"
  ),
  col_types = cols(.default = col_character()),  # read all as text, converted below
  na = c("", "NA")          # set missing values to NA
)
ride_data <- raw %>%
  mutate(
    
    #Re-code variable incentive
    incentive = case_match(incentive,
                           "No Rides"     ~ 0,
                           "1 Free Ride"  ~ 1,
                           "2 Free Rides" ~ 2),
    
    # Dates & Times (format is m/d/Y H:M, eastern time)
    across(c(start_date_time, end_date, recorded_date),
           ~ mdy_hm(.x, tz = "America/New_York")),
    # numeric fields
    across(c(progress, duration, latitude, longitude), as.numeric),
    # logical / categorical
    finished  = as.logical(as.integer(finished)),  # 1 = TRUE, 0 = FALSE
    Monday = as.factor(Monday),       #1-7 likert scale
    Tuesday = as.factor(Tuesday),     #1-7 likert scale
    Wednesday = as.factor(Wednesday), #1-7 likert scale
    Thursday = as.factor(Thursday),   #1-7 likert scale
    Friday = as.factor(Friday),       #1-7 likert scale
    Saturday = as.factor(Saturday),   #1-7 likert scale
    trip_purpose  = as.integer(trip_purpose),  #1-8 likert scale
    incentive = as.factor(incentive)
    
  )
#Remove unwatnted variables
ride_data <- ride_data %>%
  select(-c(Q_recaptcha_score, Q_recaptcha_status, Q_recaptcha_error,
            recipient_last_name, response_ID, IPaddress, phone_NO, trip_purpose_text, comments))
# using summary statistics it was noticed that IPAddress is not unique to a person, and provided no valuable information.

ride_data <- as.data.frame(ride_data)

#rename 
ride_data <- ride_data |>
  rename(end_date_time = end_date, recorded_date_time = recorded_date)

#response_ID is a per-submission unique ID
#each individual can have more than one IPaddress
#rearrange rows
ride_data <- ride_data %>%
  mutate(
    survey_date = as_date(start_date_time),     # drop time, keep the date
    start_time = hms::as_hms(start_date_time),
    end_date = as_date(end_date_time),
    end_time = hms::as_hms(end_date_time),
    recorded_date = as_date(recorded_date_time),
    recorded_time = hms::as_hms(recorded_date_time),
    day = str_extract(day, "^\\w+"),
    week_group = case_when(
      survey_date >= ymd("2026-04-06") & survey_date <= ymd("2026-04-11") ~ "Apr 6-11",
      survey_date >= ymd("2026-04-20") & survey_date <= ymd("2026-04-25") ~ "Apr 20-25",
      survey_date >= ymd("2026-05-04") & survey_date <= ymd("2026-05-09") ~ "May 4-9",
      survey_date >= ymd("2026-05-18") & survey_date <= ymd("2026-05-23") ~ "May 18-23",
      TRUE ~ NA_character_
    ),
    week_group = factor(week_group,
                        levels = c("Apr 6-11", "Apr 20-25", "May 4-9", "May 18-23"))
  ) %>% 
  arrange(week_group, desc(progress), start_date_time)  # group -> progress -> start_date_time
#remove 2 rows which had NA for week_group
ride_data <- ride_data[!is.na(ride_data$week_group), ] 
#removed 10 rows where via ID is "wilso_0"
ride_data <- ride_data |>
  mutate(unique_id   = str_extract(as.character(via), "\\d{7}")   
         #,has_survey = TRUE    # intrduce a variable to marks rows that carry a survey
  ) |>
  filter(!is.na(unique_id))

#remove unwanted rows after adjustments
ride_data <- ride_data %>%
  select(-c(start_date_time, end_date_time, recorded_date_time, via))


ride_data <- ride_data |>
  relocate(survey_date, start_time, end_date, end_time,recorded_date, recorded_time, duration, unique_id)

# convert columns day and week_group into factors
ride_data <- ride_data |>
  mutate(
    day = factor(day,
                 levels = c("Monday","Tuesday","Wednesday",
                            "Thursday","Friday","Saturday")),
    week_group = factor(week_group,
                        levels = c("Apr 6-11","Apr 20-25","May 4-9","May 18-23"))
  )

head(ride_data)
```

## Three conditions to merge on

\(1\) unique_id (in ride_data) should match with Rider_ID (in ride_full)

\(2\) Filter first_ride_of_day == TRUE (in ride_full)

\(3\) creation_date (in ride_full) should match survey_date (in ride_data)

## Data Cleaning

#### (1) Filtering ride_data to have only rows that only correspond to Request_Status == "Completed" in the ride_full data set.

There are 387 such rows corresponding to 313 individuals.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
noncomplete_ids <- ride_full |>
  filter((Request_Status != "Completed" | is.na(Request_Status)),
         first_ride_of_day == TRUE) |>
  distinct(Rider_ID, creation_date)

ride_data |>
  semi_join(noncomplete_ids,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date")) |>
  nrow()

ride_data |>
  semi_join(noncomplete_ids,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date")) |>
  distinct(unique_id) |>
  nrow()
```

Remove these 387 rows from ride_data because they correspond to rows where the ride was not completed by the individuals who started the survey.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
#remove these 387 rows, why? the ride was not completed by the individuals who started this survey
ride_data <- ride_data |>
  anti_join(noncomplete_ids,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date")) #1573 rows
nrow(ride_data)
```

After removing these rows ride_data now has 1573 rows.

#### (2) Filtering ride_full for Request_Status == "Completed"

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
ride_full <- ride_full |>
  filter(Request_Status == "Completed") 
nrow(ride_full)
```

After removing these rows ride_full now has 34,852 rows.

#### (3) Filter ride_data to have rows where unique_id (in ride_data) matches with Rider_ID (in ride_full).

There are 92 such rows corresponding to 82 individuals.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
id_mismatch = ride_data |> 
  anti_join(ride_full, by = c("unique_id" = "Rider_ID")) 
nrow(id_mismatch)
n_distinct(id_mismatch$unique_id)
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
ride_data <- ride_data |>
  semi_join(ride_full, by = c("unique_id" = "Rider_ID")) #1481 rows
nrow(ride_data)
```

After removing these rows ride_data now has 1481 rows.

#### (4) Each row in ride_data should have a distinct pair of (unique_id, survey_date).

There are only 1413 rows in ride_data with a distinct pair of both unique_id and survey_date. But the number of rows in ride_data is 1481.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# obtain distinct pairs of (unique_id, survey_date) from ride_data
surveys <- ride_data |>
  distinct(unique_id, survey_date) 
nrow(surveys) #1413 rows 
#a person-day appearing 3 times contributes 2 to this 68 (1481-1413)
```

1413 \< 1481, therefore, there are dupes in the survey data which must be removed. Dupes happen when an individual takes more than one survey in a day. There are 68 (1481-1413) such dupe surveys which must be removed.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
dupes <- ride_data |>
  group_by(unique_id, survey_date) |>
  filter(n() > 1) |>
  ungroup()  #133 dupes (this is larger than 68 because it includes the first survey taken on each day as well)
#nrow(dupes)

#Remove 68 surveys 
ride_data <- ride_data |>
  group_by(unique_id, survey_date) |>
  slice_min(start_time, n = 1, with_ties = FALSE) |>
  ungroup()

nrow(ride_data)
```

After removing these rows ride_data now has 1413 rows.

#### (5) Filter out the rows in ride_data that don't match with a row in ride_full when the 3 merging conditions are applied.

There are 155 such rows.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
ride_full_first <- ride_full |>
  filter(first_ride_of_day == TRUE)          # condition 2

ride_data |>
  anti_join(ride_full_first,
            by = c("unique_id"   = "Rider_ID",           # condition 1
                   "survey_date" = "creation_date")) |>  # condition 3
  nrow()
```

Remove these 155 surveys, they correspond to individuals who took a survey but didn't book a ride right afterwards.

Yes, removing these rows is the correct decision, since the well_being scores are particular to that specific date.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
ride_data = ride_data |>
  semi_join(ride_full_first,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date"))
nrow(ride_data)
```

After removing these rows ride_data now has 1258 rows.

#### Let's also find the number of rows in ride_data which correspond to Booking_Method == c("Agent", "Web Application") in the ride_full dataset (the 3 merging conditions are used here). It was decided to keep these surveys.

There are 9 such surveys corresponding to 8 individuals.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
agent_web_ids <- ride_full |>
  filter(Booking_Method %in% c("Agent", "Web Application"),
         first_ride_of_day == TRUE) |>
  distinct(Rider_ID, creation_date)

# ride_data rows matching on both id and date
ride_data |>
  semi_join(agent_web_ids,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date")) |>
  nrow()

# distinct individuals
ride_data |>
  semi_join(agent_web_ids,
            by = c("unique_id"   = "Rider_ID",
                   "survey_date" = "creation_date")) |>
  distinct(unique_id) |>
  nrow()
```

#### 🌟 Table summarizing the Data Cleaning carried out.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

cleaning_steps <- tribble(
  ~Description, ~rows_removed_ride_data, ~rows_removed_ride_full, ~nrow_ride_data, ~nrow_ride_full,
  "Step 0: Initial datasets before any cleaning.",
  NA,  NA,  1960L, 45190L,
  "Step 1: Filtering ride_data to have only rows that correspond to Request_Status == \"Completed\" in the ride_full dataset.",
  387L, NA,  1573L, 45190L,
  "Step 2: Filtering ride_full for Request_Status == \"Completed\".",
  NA,  10338L, 1573L, 34852L,
  "Step 3: Filter ride_data to have rows where unique_id (in ride_data) matches with Rider_ID (in ride_full).",
  92L, NA,  1481L, 34852L,
  "Step 4: Each row in ride_data should have a distinct pair of (unique_id, survey_date).",
  68L, NA,  1413L, 34852L,
  "Step 5: Filter out the rows in ride_data that don't match with a row in ride_full when the 3 merging conditions are applied.",
  155L, NA,  1258L, 34852L
)
cleaning_steps |>
  gt() |>
  tab_header(
    title = "Summary of Data Cleaning Steps",
  ) |>
  cols_label(
    Description            = "Description",
    rows_removed_ride_data = html("nrow removed from <br>ride_data"),
    rows_removed_ride_full = html("nrow removed from <br>ride_full"),
    nrow_ride_data         = html("nrow<br>ride_data"),
    nrow_ride_full         = html("nrow<br>ride_full")
  ) |>
  sub_missing(columns = everything(), missing_text = "—") |>
  cols_align(align = "center",
             columns = c(rows_removed_ride_data, rows_removed_ride_full,
                         nrow_ride_data, nrow_ride_full)) |>
  cols_align(align = "left", columns = Description) |>
  cols_width(
    Description ~ px(340),
    everything() ~ px(90)
  ) |>
  tab_options(table.font.size = px(12))
```

## Creating New Variables

### (1) Reverse Scale Saturday in ride_data

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_data <- ride_data %>%
  mutate(
    Saturday_raw = as.integer(as.character(Saturday)),
    Saturday     = factor(8L - Saturday_raw, levels = 1:7, ordered = TRUE)
  )
```

The variable *Saturday* is now reversed. The variable *Saturday_raw* has the initial variable before reversing.

### (2) Creating the variable hour in ride_data

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_data <- ride_data |>
  mutate(hour = as.numeric(start_time) %/% 3600)
```

### (3) Creating the variable well_being_score in ride_data

There are 109 rows with NA values in the well_being_score column.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_data <- ride_data %>%
  mutate(
    well_being_score = coalesce(
      as.integer(as.character(Monday)),
      as.integer(as.character(Tuesday)),
      as.integer(as.character(Wednesday)),
      as.integer(as.character(Thursday)),
      as.integer(as.character(Friday)),
      as.integer(as.character(Saturday))
    )
  )

sum(is.na(ride_data$well_being_score))
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_data %>%
  group_by(incentive) %>%
  summarise(
    n          = n(),
    wb_NA = sum(is.na(well_being_score)),
    wb_NA_rate    = mean(is.na(well_being_score))   # proportion missing
  ) %>%
  mutate(wb_NA_pct = round(100 * wb_NA_rate, 1)) %>%
  gt() %>%
  tab_header(
    title    = "Missing Well-Being Scores by Incentive Level"
  )
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

ride_data %>%
  count(finished,
        wb_is_NA = is.na(well_being_score)) %>%
  tidyr::pivot_wider(names_from  = wb_is_NA,
                     values_from = n,
                     names_prefix = "wb_NA_",
                     values_fill = 0) %>%
  gt() %>%
  tab_header(
    title    = "Survey Completion vs. Well-Being Score Missingness"
  )
```

### (4) Creating column weekly_ride_count in ride_data

Create a new data frame `pairs`. An example of what this data frame includes: If a person took 4 rides in the week before a given survey, that survey shows up as 4 rows in `pairs`.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# 1. ride log for survey takers from ride_full - individuals who didn't take even a single survey were not considered.
ride_full_ids <- ride_full |>
  filter(Rider_ID %in% ride_data$unique_id) |>
  select(unique_id = Rider_ID, ride_date = creation_date)


# 2. perform an inner_join
pairs <- ride_data |>
  inner_join(ride_full_ids, by = "unique_id", relationship = "many-to-many") |>
  mutate(days_between = as.integer(survey_date - ride_date)) |>
  filter(days_between > 0, days_between <= 7) |>
  arrange(unique_id)  

#ride_date corresponds to the creation_date in ride_full 
table(pairs$days_between) 
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# 3. one row per survey, counting how many rides fell inside each survey's lookback window.
counts <- pairs |>
  count(unique_id, survey_date, name = "weekly_ride_count") # 1171 rows

# 4. attach the counts to ride_data; surveys with no rides in the window get a count of zero.
ride_data <- ride_data |>
  left_join(counts, by = c("unique_id", "survey_date")) |>
  mutate(weekly_ride_count = coalesce(weekly_ride_count, 0L))

```

How many surveys with zero rides in weekly_ride_count were recorded?

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
sum(ride_data$weekly_ride_count == 0)
```

How many NAs are present in the weekly_ride_count column?

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
sum(is.na(ride_data$weekly_ride_count)) 
```

### (5) More Weekly Variables

Creating the following six weekly variables in ride_data:

- weekly_ride_distance

- weekly_pickup_walking_distance_ft\*

- weekly_dropoff_walking_distance_ft

- weekly_ride_duration

- weekly_waiting_time_to_pickup_minutes

- weekly_deviation_minutes\*

Columns Ride_Distance, Ride_Duration and Dropoff_Walking_Distance_ft in ride_full have NA values. Therefore, it is best to not include them in the analysis if possible. Based on the correlation between these six weekly variables (which we will be exploring), these variables were not suitable to be included in the analysis anyway.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
ride_full |>
  filter(Rider_ID %in% ride_data$unique_id) |>
  select(Ride_Distance,
         Ride_Duration,
         Pickup_Walking_Distance_ft,
         Dropoff_Walking_Distance_ft,
         Waiting_time_to_pickup_minutes,
         Deviation_minutes) |>
  summary()
```

Summary statistics of the newly created 6 weekly variables.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
#weekly ride characteristics, built with the same 7-day rule used for weekly_ride_count.
# ---------------------------------------------------------------

# 1. lean ride log for survey takers, carrying the six measures
ride_log <- ride_full |>
  filter(Rider_ID %in% ride_data$unique_id) |>
  select(unique_id = Rider_ID,
         ride_date = creation_date,
         Request_Status,
         Ride_Distance,
         Pickup_Walking_Distance_ft,
         Dropoff_Walking_Distance_ft,
         Ride_Duration,
         Waiting_time_to_pickup_minutes,
         Deviation_minutes)

# 2. one row per (survey, prior ride) pair inside the 7-day window
pairs_measures <- ride_data |>
  inner_join(ride_log, by = "unique_id", relationship = "many-to-many") |>
  mutate(days_between = as.integer(survey_date - ride_date)) |>
  filter(days_between > 0, days_between <= 7)

# 3. collapse to one row per survey
weekly_measures <- pairs_measures |>
  group_by(unique_id, survey_date) |>
  summarise(
    weekly_completed_ride_count          = n(),
    
    # totals over the week
    weekly_ride_distance                 = sum(Ride_Distance,                  na.rm = TRUE),
    weekly_pickup_walking_distance_ft    = sum(Pickup_Walking_Distance_ft,     na.rm = TRUE),
    weekly_dropoff_walking_distance_ft   = sum(Dropoff_Walking_Distance_ft,    na.rm = TRUE),
    weekly_ride_duration                 = sum(Ride_Duration,                  na.rm = TRUE),
    weekly_waiting_time_to_pickup_minutes = sum(Waiting_time_to_pickup_minutes, na.rm = TRUE),
    weekly_deviation_minutes             = sum(Deviation_minutes,              na.rm = TRUE),
    .groups = "drop"
  )

# attach to ride_data
# Totals: a survey with no completed rides in the window gets 0.
ride_data <- ride_data |>
  left_join(weekly_measures, by = c("unique_id", "survey_date")) |>
  mutate(
    weekly_completed_ride_count = coalesce(weekly_completed_ride_count, 0L),
    across(c(weekly_ride_distance,
             weekly_pickup_walking_distance_ft,
             weekly_dropoff_walking_distance_ft,
             weekly_ride_duration,
             weekly_waiting_time_to_pickup_minutes, #is median more suited ? No
             weekly_deviation_minutes #is median more suited ? No
    ),
    ~ coalesce(.x, 0))
  )

# 5. check
summary(select(ride_data, weekly_ride_distance, weekly_ride_duration))
summary(select(ride_data, weekly_pickup_walking_distance_ft, weekly_dropoff_walking_distance_ft))
summary(select(ride_data, weekly_waiting_time_to_pickup_minutes, weekly_deviation_minutes))
```

Exploring collinearity between these six variables to determine which to include in the analysis.

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.width=20, fig.height=15}
# ---------------------------------------------------------------
# exploring the collinearity beteen  the weekly ride variables
# ---------------------------------------------------------------

collinearity_vars <- ride_data |>
  select(weekly_ride_count,
         weekly_ride_distance,
         weekly_pickup_walking_distance_ft,
         weekly_dropoff_walking_distance_ft,
         weekly_ride_duration,
         weekly_waiting_time_to_pickup_minutes,
         weekly_deviation_minutes)

# Pearson correlation matrix as a table
#round(cor(collinearity_vars, method = "pearson", use = "complete.obs"), 3)

# scatter plot matrix

scatter_matrix <- ggpairs(
  collinearity_vars,
  lower = list(continuous = wrap("smooth", alpha = 0.25, size = 0.5,
                                 colour = "grey40",
                                 se = FALSE, method = "loess")),
  diag  = list(continuous = wrap("densityDiag", fill = "steelblue", alpha = 0.5)),
  upper = list(continuous = wrap("cor", method = "pearson", size = 12)),
  progress = FALSE
) +
  labs(title = "Collinearity among weekly ride characteristics",
       subtitle = "Pearson correlations; 7-day lookback totals") +
  theme_bw(base_size = 20) +
  theme(strip.text = element_text(size = 10),
        axis.text  = element_text(size = 15))

scatter_matrix
```

To avoid collinearity within the statistical model, we decided to include only the following weekly variables:

- weekly_ride_count

- ONE walking measure: weekly_pickup_walking_distance_ft

- weekly_deviation_minutes.

## Summary Statistics

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# No. of incentives per incentive category
incentive_summary <- ride_data %>%
  mutate(incentive = factor(incentive, levels = c(0, 1, 2),
                            labels = c("No Rides", "1 Free Ride", "2 Free Rides"))) %>%
  count(incentive, name = "No_of_incentives") %>%
  mutate(percent = round(No_of_incentives / sum(No_of_incentives) * 100, 1)) %>%
  rename(`Incentive`   = incentive,
         `Submissions` = No_of_incentives,
         `Percent`     = percent)

incentive_summary %>% gt()
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
frequency_summary <- ride_data %>%
  count(unique_id, name = "n_sub") %>%
  mutate(times = case_when(
    n_sub == 1 ~ "Once",          n_sub == 2 ~ "Twice",
    n_sub == 3 ~ "Three times",   n_sub == 4 ~ "Four times",
    n_sub == 5 ~ "Five times",    n_sub == 6 ~ "Six times",
    n_sub == 7 ~ "Seven times",   n_sub == 8 ~ "Eight times",
    n_sub == 9 ~ "Nine times",    n_sub == 10 ~ "Ten times",
    n_sub  > 10 ~ "More than 10 times"
  )) %>%
  count(times, name = "participants") %>%
  mutate(times = factor(times, levels = c("Once","Twice","Three times","Four times",
                                          "Five times","Six times","Seven times",
                                          "Eight times","Nine times","Ten times",
                                          "More than 10 times"))) %>%
  arrange(times) %>%
  mutate(percent = round(participants / sum(participants) * 100, 1)) %>%
  rename(`Times surveyed` = times,
         `Participants`   = participants,
         `Percent`        = percent)

frequency_summary %>% gt()
```

## 1.0 Logistic Regression Model

### 1.1 Why a individual specific random intercept was not included in the Logistic Model.

$$\begin{aligned}
Y_{ij}| u_j &\sim \text{Bernoulli}(\pi_{ij}(u_j)), \quad Y_{ij} = 1 \text{ when survey } i \text{ is completed by individual } j \\[6pt]
\log\!\left(\frac{\pi_{ij}}{1 - \pi_{ij}}\right) &= (\beta_0 + u_j) + \beta_1\,\text{OneFreeRide}_{ij} + \beta_2\,\text{TwoFreeRides}_{ij} + \beta_3\,\text{day}_{ij} \\[6pt]
u_j &\overset{\mathrm{iid}}{\sim} N(0, \sigma_u^2)
\end{aligned}$$

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

model_df <- ride_data %>%
  mutate(incentive = factor(incentive, levels = c(0, 1, 2),
                            labels = c("No Rides", "1 Free Ride", "2 Free Rides")))

# finished (T/F) ~ incentive + day + random intercept per individual 
logistic_model <- glmer(
  finished ~ incentive + day + (1 | unique_id),
  family = binomial,
  data   = model_df,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e6))
)


summary(logistic_model)
VarCorr(logistic_model)
```

- **Intercept (8.44):** A typical rider (random intercept = 0) offered No Rides on a Friday (the reference day), has predicted log odds of completing the survey of 8.44. probability of completion: e\^8.44 / (1 + e\^8.44) ≈ 0.9997

- **Random intercept (** $\hat{\sigma}_u$ **= 11.727, variance = 137.52):** This captures how much an individual rider's baseline log odds of completing departs from the overall intercept. A SD of 11.727 on the logit scale is extremely large, it corresponds to fitted individual completion probabilities pinned at essentially 0 or 1.

$$L_j( \beta, \sigma_u^2) = f(y_j) = \int_{- \infty}^{\infty} f(y_j,u_j) du_j =  \int_{- \infty}^{\infty} f(y_j|u_j)f(u_j)du_j \\ $$

$$\begin{aligned}
L_j(\boldsymbol\beta, \sigma_u^2)
&= \int_{-\infty}^{\infty}
\left[ \prod_{i=1}^{n_j} \pi_{ij}(u_j)^{\,y_{ij}}\bigl\{1-\pi_{ij}(u_j)\bigr\}^{\,1-y_{ij}} \right]
\frac{1}{\sqrt{2\pi}\,\sigma_u}\,
\exp\!\left(-\frac{u_j^{2}}{2\sigma_u^{2}}\right) du_j
\end{aligned}$$

$$\ell(\underline{\beta},\sigma_u^{2}) = \sum_{j=1}^G \ell_j(\underline{\beta},\sigma_u^{2}) \\$$

$$ \pi_{ij}(u_j) = \frac{e^{(\beta_0+u_j) + \underline{X}^T_{ij} \underline{\beta}}}{1+ e^{(\beta_0+u_j) + \underline{X}^T_{ij} \underline{\beta}}} \\ $$

Most riders opted to complete the survey, so on most surveys $y_{ij}=1$ . When a rider decides to complete every survey they take, the conditional likelihood $\prod_{i=1}^{n_j}\pi_{ij}(u_j)$ is strcktly increasing in $u_j$ with no maximum, so $\hat u_j \rightarrow +\infty$ and in this data set it is so for many riders :

$$ \prod_{i=1}^{n_j}\pi_{ij}(u_j) \;\xrightarrow[\;u_j\to+\infty\;]{}\; 1, \qquad \hat u_j = +\infty, $$

Similarly, when a rider decides to not complete every survey they started, $\prod_{i=1}^{n_j}\pi_{ij}(u_j) \rightarrow 0$ and $\hat u_j \rightarrow -\infty$. Because $\sigma_u^2$ measures the spread of these rider intercepts, driving a large share of them toward $\pm\infty$ makes the log-likelihood $\ell(\underline{\beta}, \sigma^2_u)$ increase without bound in $\sigma_u^2$,

$$ 
\frac{\partial}{\partial \sigma_u^{2}}\,\ell(\underline{\beta},\sigma_u^{2}) > 0 $$

so no finite maximum exists over the explored range, for the above expression. The reported $\hat\sigma_u = 11.727$ (variance $137.52$) is because it stops at a numerical tolerance rather than an estimate of genuine between-rider heterogeneity.

An implied intra-class correlation of 0.977 asserts that almost all variation in completion is within riders and almost none between; and the fitted individual probabilities are pinned at the boundaries 1 or 0.

$$\hat{ICC} = \frac{\hat{\sigma}_u^2}{\hat{\sigma}_u^2 + \frac{\pi^2}{3}} = \frac{137.52}{137.52 + 3.29} \approx 0.977$$

The inflated $\hat{\sigma}_u$ and the implied $\hat{ICC} \approx 0.977$ imply cluster level separation; because nearly every rider either completed all or none of their surveys, each rider's intercept is driven toward $\pm\infty$, and the variance component $\sigma_u^2$ that summarises their spread is inflated to accommodate these boundary values rather than measuring genuine between-rider heterogeneity.

It is evident from the table below that individuals who had taken the survey multiple times were more likely to complete it.

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}

completion_by_frequency <- ride_data %>%
  # how many surveys did each rider receive?
  add_count(unique_id, name = "n_sub") %>%
  mutate(times = case_when(
    n_sub == 1 ~ "Once",          n_sub == 2 ~ "Twice",
    n_sub == 3 ~ "Three times",   n_sub == 4 ~ "Four times",
    n_sub == 5 ~ "Five times",    n_sub == 6 ~ "Six times",
    n_sub == 7 ~ "Seven times",   n_sub == 8 ~ "Eight times",
    n_sub == 9 ~ "Nine times",    n_sub == 10 ~ "Ten times",
    n_sub  > 10 ~ "More than 10 times"
  ),
  times = factor(times, levels = c("Once","Twice","Three times","Four times",
                                   "Five times","Six times","Seven times",
                                   "Eight times","Nine times","Ten times",
                                   "More than 10 times")),
  status = factor(if_else(finished, "Completed", "Not completed"),
                  levels = c("Completed", "Not completed"))) %>%
  # count surveys in each (times, status) cell
  count(times, status, name = "n", .drop = FALSE) %>%
  # make sure every times x status combination appears, even with zero surveys
  tidyr::complete(times, status, fill = list(n = 0L)) %>%
  # convert to % within times
  group_by(times) %>%
  mutate(total   = sum(n),
         percent = if_else(total > 0, 100 * n / total, 0)) %>%
  ungroup() %>%
  select(-total)

ggplot(completion_by_frequency, aes(times, percent, fill = status)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3) +
  scale_y_continuous(breaks = seq(0, 100, by = 10),
                     labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = c(0, 0.12))) +
  scale_fill_manual(values = c("Completed" = "steelblue",
                               "Not completed" = "firebrick"),
                    name = NULL) +
  labs(title = "Survey completion by how often a rider was surveyed",
       subtitle = "Bars give the percentage within each category; labels give the number of surveys",
       x = "Times surveyed", y = "Percentage of surveys") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top")
```

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}

plot_df <- ride_data %>%
  filter(!is.na(incentive), !is.na(finished)) %>%
  mutate(
    incentive = factor(incentive, levels = c(0, 1, 2),
                       labels = c("No Rides", "1 Free Ride", "2 Free Rides")),
    finished  = factor(finished, levels = c(FALSE, TRUE),
                       labels = c("Not finished", "Finished"))
  )

# counts + percentage within each incentive group
summary_df <- plot_df %>%
  count(incentive, finished, name = "n") %>%
  group_by(incentive) %>%                       # denominator = total per incentive
  mutate(pct = n / sum(n) * 100) %>%
  ungroup()

ggplot(summary_df, aes(x = finished, y = pct, fill = finished)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.3) +
  facet_wrap(~ incentive, nrow = 1) +
  scale_y_continuous(limits = c(0, 100),
                     breaks = seq(0, 100, by = 20),
                     labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = c(0, 0.12))) +  # headroom for labels
  labs(x = NULL, y = "Percentage of surveys") +
  theme_bw()
```

### 1.2 Logistic Model that works

Resolved to fit the logistic model without a random intercept.

Drawback: glm() treats all 1258 surveys as unrelated, but they come from 648 riders, so repeat respondents are counted as independent evidence when they are not.

```{r, message=FALSE, warning=FALSE, comment=NA}

model_df <- ride_data %>%
  mutate(incentive = factor(incentive, levels = c(0, 1, 2),
                            labels = c("No Rides", "1 Free Ride", "2 Free Rides")))

glm_base1 <- glm(finished ~ incentive  + week_group , family = binomial, data = model_df)

glm_base2 <- glm(finished ~ incentive + hour + day + week_group , family = binomial, data = model_df)

glm_ride_vol1 <- glm(finished ~   weekly_ride_count + incentive + week_group, family = binomial, data = model_df) 


# finished ~   weekly_ride_count * incentive + week_group
#because 2 rides were only offered in the last two weeks; finished ~   weekly_ride_count + incentive*week_group

glm_ride_vol2 <- glm(finished ~ incentive + weekly_ride_count + hour + day + week_group, family = binomial, data = model_df)


glm_all <- glm(finished ~ incentive + weekly_ride_count + hour + day + week_group + weekly_deviation_minutes + weekly_pickup_walking_distance_ft, family = binomial, data = model_df)
  
BIC(glm_base1, glm_base2, glm_ride_vol1, glm_ride_vol2, glm_all)
AIC(glm_base1, glm_base2, glm_ride_vol1, glm_ride_vol2, glm_all)
```

Decided on the model $\textit{glm\_ride\_vol1}$ since it has lowest AIC and BIC values:

$$\textit{finished} \sim \textit{incentive + weekly\_ride\_count + week\_group} $$

$$Y_{ij} \sim \operatorname{Bernoulli}\left(\pi_{ij}\right), \qquad Y_{ij} = 1 \text{ if survey } i \text{ is completed by individual } j.$$

$$\log\!\left(\frac{\pi_{ij}}{1 - \pi_{ij}}\right) = (\beta_0 ) + \beta_1\,\text{OneFreeRide}_{ij} + \beta_2\,\text{TwoFreeRides}_{ij} + \beta_3\,\text{WeeklyRideCount}_{ij} + \\ 
\beta_4\,\text{Apr20-25}  + \beta_5\,\text{May4-9} + \beta_6\,\text{May18-23}$$

$hour$ and $day$ variables were not significant, and were not included in this logistic model.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
summary(glm_ride_vol1) #CENTER AT mean/median rides
```

The drop from null to residual is the deviance explained by the predictors:

- 764.79−726.89 = 37.9 on 1257−1251=6 degrees of freedom

Under the null hypothesis that all seven slopes are zero, this difference follows a $\chi^2_{6}$​ distribution:

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
pchisq(37.9, df = 6, lower.tail = FALSE)   
```

This p-value is the probability of observing a deviance drop *at least as large as* 37.9, **assuming all six slopes are truly zero**. This null hypothesis is rejected.

$$p = P(\chi^2_6 \geq 37.9 | H_0 \ true) \approx 0 $$

### 1.3 Using the sandwich estimator

Two assumptions about how the surveys are related give two different variance estimators for the same coefficients.

**Assumption 1:** all 1441 surveys are independent, then the variance of the coefficients takes the following form.

$$\hat{Var}(\hat{\beta}) = (X^TWX)^{-1}, \ W = diag(\hat{\pi_i}(1-\hat{\pi}_i))$$

**Assumption 2:** clustering based on rider $j$ using the sandwich estimator. In this estimator rider $j$ enters only through the clustering of the residuals and not through a parameter, so the coefficients are population averaged rather than subject specific.

Both forms are obtained from the same derivation, which is set out below.

**Deriving the variance of the coefficients:**

The estimator $\hat{\boldsymbol\beta}$ solves the logistic score equations, which set the total score to zero:

$$\mathbf{U}(\boldsymbol\beta) = \sum_{i=1}^{n} (y_i - \pi_i)\,\mathbf{x}_i = \mathbf{0}.$$

First order Taylor expansion of $\mathbf{U}(\hat{\boldsymbol\beta})$ about the true value $\boldsymbol\beta_0$ :

$$\begin{aligned}
\mathbf{0} = \mathbf{U}(\hat{\boldsymbol\beta})
&\approx \mathbf{U}(\boldsymbol\beta_0) + \frac{\partial \mathbf{U}}{\partial \boldsymbol\beta^\top}\Big|_{\boldsymbol\beta_0} (\hat{\boldsymbol\beta} - \boldsymbol\beta_0), \\[4pt]
\Rightarrow\quad
(\hat{\boldsymbol\beta} - \boldsymbol\beta_0)
&\approx \underbrace{\Bigl(-\tfrac{\partial \mathbf{U}}{\partial \boldsymbol\beta^\top}\Bigr)^{-1}}_{\text{bread}} \, \mathbf{U}(\boldsymbol\beta_0),
\end{aligned}$$

then take the variance of both sides;

$$\operatorname{Var}(\hat{\boldsymbol\beta})
\approx
\Bigl(-\tfrac{\partial \mathbf{U}}{\partial \boldsymbol\beta^\top}\Bigr)^{-1}
\operatorname{Var}\!\bigl(\mathbf{U}(\boldsymbol\beta_0)\bigr)
\Bigl(-\tfrac{\partial \mathbf{U}}{\partial \boldsymbol\beta^\top}\Bigr)^{-1}.$$

- **Bread :**

The bread is the negative derivative of the score which for logistic regression is the information matrix:

$$-\frac{\partial \mathbf{U}}{\partial \boldsymbol\beta^\top}
= \sum_{i=1}^{n} \pi_i(1-\pi_i)\,\mathbf{x}_i \mathbf{x}_i^\top
= X^\top W X,
\qquad W = \operatorname{diag}\!\bigl(\hat\pi_i(1-\hat\pi_i)\bigr).$$

- **Meat :**

The meat is $\operatorname{Var}(\mathbf{U})$, and this is the step where either Assumption 1 or Assumption 2 can be used.

In Assumption 2, we write the score as a sum of the per-rider contributions $\mathbf{s}_j = \sum_{i \in j}(y_{ij}-\pi_{ij})\mathbf{x}_{ij}$, where $\mathbf{s}_j$ is the sum of the score contributions by rider $j$.

$$\mathbf{U} = \sum_{j=1}^{G} \mathbf{s}_j,
\qquad
\operatorname{Var}(\mathbf{U}) = \sum_{j=1}^{G}\operatorname{Var}(\mathbf{s}_j)
\;\;\text{if riders are mutually independent.}$$

Two choices for estimating this meat give the two variance formulas:

$$\begin{aligned}
\text{Assumption 1: treat all } n \text{ surveys as independent:}\quad
&\operatorname{Var}(\mathbf{U}) = \sum_{i=1}^{n}(y_i-\pi_i)^2\,\mathbf{x}_i\mathbf{x}_i^\top
\;\to\; X^\top W X, \\[4pt]
\text{Assumption 2: treat only the } G \text{ riders as independent:}\quad
&\widehat{\operatorname{Var}}(\mathbf{U}) = \sum_{j=1}^{G} \mathbf{s}_j \mathbf{s}_j^\top .
\end{aligned}$$

- **The two estimators :**

Under the first assumption the extimated variance of the coefficients collapses back to the bread and become; $\hat{\operatorname{Var}}(\hat{\boldsymbol\beta}) = (X^\top W X)^{-1}$.

The second cluster-based assumption, yields the cluster-robust estimator for $G=648$ riders,

$$\hat{\operatorname{Var}}(\hat{\boldsymbol\beta})
= (X^\top W X)^{-1}\!\left[\sum_{j=1}^{648}\mathbf{s}_j\mathbf{s}_j^\top\right]\!(X^\top W X)^{-1},
\qquad
\mathbf{s}_j = \sum_{i\in j}(y_{ij}-\hat\pi_{ij})\,\mathbf{x}_{ij}.$$

The only difference between these two approaches is how the residuals are aggregated within the meat of the sandwich estimator. Under the first assumption, all rows are considered independent, so the residuals are aggregated by squaring them survey by survey. Under the second assumption, the residuals are first summed within each rider and then squared.

The essential point is in that inner sum. Residuals are added up within each rider before being squared, so if a rider's residuals move together the meat inflates and the standard errors widen. If residuals within a rider are unrelated, the meat is close to the independence version and nothing much changes.

On evaluating the coefficients and This cluster-robust fit is empirical evidence that the within-rider correlation in completion is negligible.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

coeftest(glm_ride_vol1, vcov = vcovCL, cluster = ~ unique_id)
```

**Intercept (1.5651):** A rider offered No Rides in the reference week (Apr 6-11) with no rides in the preceding week has predicted log-odds of completing the survey of 1.57. Probability of completion: $\frac{e^{1.57}}{1 + e^{1.57}} ≈ 0.83$.

**1 Free Ride (0.5070):** Holding week group and weekly ride count fixed, being offered 1 free ride rather than none raises the log-odds of completion by 0.51. On the odds scale, $e^{0.51} ≈ 1.66$, so the odds of completing are roughly 1.7 times higher than under No Rides (p = 0.0130). On the probability scale this moves the reference rider from 0.83 ( $\frac{ e^{(1.5651)} }{{1+e^{(1.5651)}} }$ ) to about 0.89 ( $\frac{e^{(1.5651+0.5070)}}{1+e^{(1.5651+0.5070)}}$ ).

**2 Free Rides (1.2935):** Relative to No Rides, the log-odds rise by 1.29. On the odds scale, $e^{1.29} ≈ 3.65$, so the odds of completing are roughly 3.6 times higher than under No Rides (p = 0.0035). In probability terms, from 0.83 to about 0.94 ( $\frac{e^{(1.5651+1.2935)}}{1 + e^{(1.5651+1.2935)}} ≈ 0.83$ ).

The monotone ordering holds (0 \< 0.51 \< 1.29), and both contrasts remain significant after accounting for repeated responses by the same rider (which is the result we wanted), since it means the incentive effect is not an artifact of a handful of frequent respondents. The step from one free ride to two is larger than the step from none to one: the odds ratio for two rides is about 2.2 times that for one, so the return to the second ride is greater than the first rather than diminishing.

**Weekly ride count (0.0648):** Each additional ride taken in the seven days before the survey raises the log-odds of completion by 0.065, an odds ratio of $e^{0.065} ≈ 1.07$ per ride (p = 0.0246). The per-ride effect is small but accumulates: five additional rides multiply the odds by about 1.38 and ten by about 1.91. For the reference rider this corresponds to a rise in completion probability from 0.83 at zero recent rides to about 0.87 at five rides and 0.90 at ten. Riders who use the service more frequently are therefore more likely to finish the survey, which is consistent with greater engagement with the app.

**Week group (reference = Apr 6-11):** May 18-23 is the only week that differs significantly from the reference (1.1183, OR ≈ 3.06, p = 0.0075), indicating higher completion late in the study. Apr 20-25 is indistinguishable from the reference (0.0105, OR ≈ 1.01, p = 0.966), and May 4-9 is slightly lower but not significant (−0.2128, OR ≈ 0.81, p = 0.456). Completion is therefore broadly stable across the first three survey weeks, with a rise in the final week.

Furthermore, it is evident that this Ride data is cluster robust.

### 1.4 Evaluating Goodness of fit

Goodness of fit asks: given this data set, is the fitted model adequate or does the model systematically mis-predict/under-perform in some region?

This is an in-sample question, exactly like residual plots in linear regression.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
library(patchwork)
model_df <- model_df %>% mutate(phat = fitted(glm_ride_vol1))
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE, fig.width=10, fig.height=5}

p_cal <- model_df %>%
  mutate(bin = ntile(phat, 10)) %>%
  group_by(bin) %>%
  summarise(n = n(), predicted = mean(phat), observed = mean(finished),
            se = sqrt(observed * (1 - observed) / n), .groups = "drop") %>%
  ggplot(aes(predicted, observed)) +
  geom_abline(linetype = "dashed", colour = "grey50") +
  geom_errorbar(aes(ymin = observed - 1.96*se, ymax = observed + 1.96*se),
                width = 0, colour = "grey60") +
  geom_point(aes(size = n), colour = "steelblue") +
  scale_size_continuous(guide = "none") +
  labs(title = "Calibration: predicted vs observed completion",
       subtitle = "Points on the dashed line indicate accurate predictions",
       x = "Mean predicted probability (decile)", y = "Observed proportion completed") +
  coord_equal() + theme_bw()
p_cal
```

Each survey $i=1,2,...,n$ has a fitted probability of completion from the logistic model:

$$ \hat{\pi}_i = \frac{\exp\left(\mathbf{x}_i^\top \hat{\boldsymbol\beta}\right)}{1 + \exp\left(\mathbf{x}_i^\top \hat{\boldsymbol\beta}\right)}, \\
\qquad \mathbf{x}_i^\top \hat{\boldsymbol\beta} = \hat\beta_0 + \hat\beta_1 \text{OneFreeRide}_{ij} + \hat\beta_2 \text{TwoFreeRides}_{ij} + \hat\beta_3\,\text{WeeklyRideCount}_{ij} +\\
\hat\beta_4\,\text{Apr20-25}  + \hat\beta_5\,\text{May4-9} + \hat\beta_6\,\text{May18-23} $$

The observations are ordered by $\hat\pi_i$​ and split at the sample deciles, giving $g=1, \dots, 10$ groups.

$B_1, \dots, B_{10}$ of approximately equal size. Writing $\hat\pi_{(1)} \le \cdots \le \hat\pi_{(n)}$ for the ordered fitted values, group $g$ collects

$$B_g = \left\{ i : \hat{\pi}_i \in \left(q_{(g-1)/10},\, q_{g/10}\right] \right\},
\qquad
q_\alpha = \text{the } \alpha \text{ quantile of } 
\{\hat{\pi}_1, \dots, \hat{\pi}_n\}.$$\

with $n_g = |B_g|$ the number of surveys in group $g$.

For each group, the horizontal coordinate is the mean fitted probability and the vertical coordinate is the observed completion proportion:

$$\bar{\pi}_g = \frac{1}{n_g}\sum_{i \in B_g} \hat{\pi}_i, \qquad
\bar{y}_g = \frac{1}{n_g}\sum_{i \in B_g} y_i = \frac{O_g}{n_g}$$

where $y_i \in \{0,1\}$ indicates completion and $\sum_{i \in B_g} y_i$ is the observed number of completers. The plot shows the pairs $(\bar\pi_g, \bar y_g)$ for g = 1,...,10 against the identity line $y=x$. Perfect calibration means:

$E[Y_i​ | \hat{π}_i​=p]=p ⟹ \bar{y}_{​g}​  \approx \bar{\pi}_g​ \text{ for all g}$

Treating the $n_g$ outcomes within a group as Bernoulli, the standard error of the observed proportion is:

$$\widehat{\text{SE}}(\bar{y}_g) = \sqrt{\frac{\bar{y}_g\left(1 - \bar{y}_g\right)}{n_g}}, \qquad
\text{The 95\% interval: } \ \bar{y}_g \pm 1.96 \cdot \widehat{\text{SE}}(\bar{y}_g)$$

Why some points are larger than others in the graph ?

Because of ties in the fitted probabilities, many riders share exactly the same covariate pattern and exactly the same $\hat{\pi}_i$ . A decile boundary can't split a tied block cleanly; all rows with an identical fitted value land together; so some bins absorb a large block of tied riders and swell, while adjacent bins are left with fewer.

## 2.0 Weighted Models - IPW for non-response on finished

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}

library(dplyr)
library(tidyr)
library(ggplot2)

day_levels <- c("Monday: overall health", "Tuesday: vigor", "Wednesday: physical health",
                "Thursday: mental health", "Friday: sleep quality", "Saturday: loneliness")

# shared lookup so both branches produce identical sector labels
day_labels <- c(Monday    = "Monday: overall health",
                Tuesday   = "Tuesday: vigor",
                Wednesday = "Wednesday: physical health",
                Thursday  = "Thursday: mental health",
                Friday    = "Friday: sleep quality",
                Saturday  = "Saturday: loneliness")

# seven graded shades + grey for missing (Missing listed first)
my_colors <- c("Missing" = "grey55",
               "1" = "#fdd49e", "2" = "#fdbb84", "3" = "#f19c7c",
               "4" = "#fc8d59", "5" = "#e34a33", "6" = "#b30000",
               "7" = "#7f0000")

# 1. answered surveys: one row per (week, weekday, Likert score)
answered <- ride_data |>
  mutate(across(c(Monday, Tuesday, Wednesday, Thursday, Friday, Saturday),
                ~ as.integer(as.character(.x)))) |>
  select(week_group, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) |>
  pivot_longer(-week_group, names_to = "day_cat", values_to = "score") |>
  filter(!is.na(score)) |>
  mutate(day_cat = unname(day_labels[day_cat])) |>
  count(week_group, day_cat, score, name = "count")

# 2. surveys with no well-being score, attributed to their own weekday
missing <- ride_data |>
  filter(is.na(well_being_score), !is.na(day)) |>
  count(week_group, day, name = "count") |>
  mutate(day_cat = unname(day_labels[as.character(day)]),
         score   = NA_integer_) |>
  select(week_group, day_cat, score, count)

# 3. combine; ensure every week x day sector exists
radial_df <- bind_rows(answered, missing) |>
  mutate(
    day_cat = factor(day_cat, levels = day_levels),
    score_f = factor(if_else(is.na(score), "Missing", as.character(score)),
                     levels = c("Missing", as.character(1:7)))   # Missing before 1
  ) |>
  complete(week_group, day_cat, fill = list(count = 0L))

# 4. radial gridline labels — max is the tallest single bar
max_r    <- max(radial_df$count, na.rm = TRUE)
r_breaks <- pretty(c(0, max_r), n = 8)
r_breaks <- r_breaks[r_breaks > 0 & r_breaks <= max_r]

label_df <- data.frame(x = 0.5, count = r_breaks, label = as.character(r_breaks))

# 5. plot
ggplot(radial_df, aes(x = day_cat, y = count, fill = score_f)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0.05),
           width = 0.9) +                                        # individual bars
  geom_text(data = label_df,
            aes(x = x, y = count, label = label),
            inherit.aes = FALSE, size = 2.6, colour = "grey30",
            hjust = 0.5, vjust = -0.3) +
  coord_polar(start = 0, clip = "off") +
  facet_wrap(~ week_group, nrow = 2) +
  scale_y_continuous(limits = c(-max_r * 0.18, max_r * 1.05),
                     breaks = r_breaks, expand = c(0, 0)) +
  scale_x_discrete(labels = function(x) gsub(": ", ":\n", x)) +   # wrap onto two lines
  scale_fill_manual(values = my_colors, name = "Well-being score",
                    na.translate = FALSE) +
  labs(
    title    = "Breakdown of the well-being score by day and study week",
    subtitle = "The radius provides the number of surveys belonging to each category; within each day, the grey bar counts surveys with no well-being response.",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y      = element_blank(),
    axis.ticks.y     = element_blank(),
    axis.text.x      = element_text(size = 10, colour = "grey20", lineheight = 0.9),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    panel.spacing    = unit(2.5, "lines"),          # room between the four panels
    plot.margin      = margin(10, 40, 10, 40),      # room at the left/right edges
    strip.text       = element_text(face = "bold", size = 12),
    legend.position  = "bottom",
    legend.key.size  = unit(0.5, "cm"),
    legend.text      = element_text(size = 10),
    legend.title     = element_text(size = 11)
  )
```

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

# ipw for the well-being ~ microtransit use association

library(dplyr)
library(ordinal)      # clmm

survey_weeks <- c("Apr 6-11", "Apr 20-25", "May 4-9", "May 18-23")

# returns the day of the week
weekday_of <- function(d) {
  factor(c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")[
    as.integer(format(d, "%u"))],
    levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
}

#eg: dd=as.Date("2026-08-10"); weekday_of(dd)

# ---------------------------------------------------------------
# 1st step: create "riders" dataframe, which includes all the riders who could have seen the
#           pop-up to take the survey. So this dataset includes the individuals who decided to #           not click on the survey as well. 
#----------------------------------------------------------------

riders <- ride_full |>
  filter(first_ride_of_day,
         Request_Status == "Completed",
    #Booking_Method == "Mobile Application", - no need because used the app to take the survey
         study_week %in% survey_weeks) |>
  transmute(unique_id   = Rider_ID,       
            survey_date = creation_date,  
            week_group  = factor(study_week, levels = survey_weeks), 
            day         = weekday_of(creation_date),
            hour        = as.numeric(creation_time) %/% 3600) 

# U = vector of microtransit utilization (rides in past week)
# ride_data already has weekly_ride_count, but only for survey takers.
# The click model needs it for riders who never clicked, so it is rebuilt here from "riders" dataframe using the same rule: (0 < gap <= 7)

lookback <- riders |> 
  dplyr::select(unique_id, survey_date) |> 
  inner_join(ride_full |> 
             transmute(unique_id = Rider_ID, ride_date = creation_date), 
             by = "unique_id", relationship = "many-to-many") |> 
  mutate(gap = as.integer(survey_date - ride_date)) |> 
  filter(gap > 0, gap <= 7) |> 
  count(unique_id, survey_date, name = "weekly_ride_count")
# colnames of lookback: unique_id, survey_date, weekly_ride_count
# nrow(lookback) = 8526

riders <- riders |>
  left_join(lookback, by = c("unique_id", "survey_date")) |>
  mutate(weekly_ride_count = coalesce(weekly_ride_count, 0L))

# ---------------------------------------------------------------
# 2nd step: Attach C, F, I and the outcome from ride_data.
# C: click = 1, binary indicator of clicking on the pop-up to see more about the survey
# F: binary indicator of finishing the survey
# I: three-level categorical incentive
# Saturday is already reversed.
# ---------------------------------------------------------------
survey_status <- ride_data |>
  transmute(unique_id, survey_date,
        click     = 1L,
        complete  = as.integer(!is.na(well_being_score)), #indicator: complete wellbeing (0/1)
        incentive = factor(incentive, levels = c(0, 1, 2)),
        wb        = well_being_score)

#match the survey to the nearest date, such that it is within the relevant week_group

riders <- riders |>
  left_join(survey_status, by = c("unique_id", "survey_date")) |> 
#since its a left join all rows of dataframe "riders" are retained, additionally columns;
#"click","complete","incentive" and "wb" from "survey_status" dataset are included. 
  mutate(click    = coalesce(click, 0L), #non-clickers are set to zero
         complete = if_else(click == 0L, 0L, coalesce(complete, 0L)
                            )
         ) 
         #for click=0, corresponding complete is set to zero

table(riders$click, riders$complete)  #column = click(0/1)

```

Number of rows in ride_data is 1258, this corresponds to the summation 109+1149 (where click=1).

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

# ---------------------------------------------------------------
# step 4 - Model 1 the click model, fit on ALL riders in the 4 week survey period
#   who had the opportunity to click on the survey.
#   P(C = 1 | R, U).  R: vector of rider characteristics (day, hour),
#   U: vector of microtransit utilization (weekly_ride_count)
# Incentive is deliberately EXCLUDED since riders cannot see it before clicking,
# so P(C = 1 | I) = P(C = 1)
# ---------------------------------------------------------------

fit_click <- glm(click ~ weekly_ride_count + hour + day + week_group,
                 family = binomial, data = riders)

# cluster robust inference (population-averaged, clustered on rider)
coeftest(fit_click, vcov = vcovCL, cluster = ~ unique_id)

# joint test of the day effect under the cluster-robust variance
waldtest(fit_click,
         update(fit_click, . ~ . - day),
         vcov = vcovCL(fit_click, cluster = ~ unique_id),
         test = "Chisq")

# joint test of the week group effect
waldtest(fit_click,
         update(fit_click, . ~ . - week_group),
         vcov = vcovCL(fit_click, cluster = ~ unique_id),
         test = "Chisq")

riders$prob_click <- predict(fit_click, type = "response")

```

There are 10,085 rows in data frame rider:

- These come from 2,060 riders (median of 2 rides per rider and maximum of 23 rides per rider).

- 1258 rows from this data set correspond to the surveys taken (no. of rows in ride_data is 1258).

When using the function glm() the standard errors treat repeated rides from the same rider as independent. Therefore, similar to section 1.3 we make use of cluster-robust standard errors on unique_id.

**Intercept (−1.9004):** A rider on a Monday in the reference week (Apr 6-11), with no rides in the preceding week, has predicted log-odds of clicking of −1.90, a probability of about 0.13. Since *hour* = 0 falls outside the service window (05:00 to 18:00), the more meaningful anchor is the median hour of 09:00, giving a clicking probability of about 0.17 ( $plogis(−1.9004+9×0.0368)$ ); the probabilities below use this reference.

**Hour (0.0368):** Each hour later in the day raises the log-odds of clicking by 0.037, an odds ratio of e\^0.037 ≈ 1.04 per hour (p = 0.0019). Across the 13-hour service day the odds multiply by about 1.61, moving the reference rider from roughly 0.15 at 05:00 to 0.23 at 18:00.

**Day (reference = Monday):** Clicking declines steadily through the week. Tuesday is indistinguishable from Monday (−0.1033, OR ≈ 0.90, p = 0.252), while Wednesday (−0.2187, OR ≈ 0.80, p = 0.0224), Thursday (−0.2812, OR ≈ 0.75, p = 0.0027), Friday (−0.4552, OR ≈ 0.63, p = 7.4 × 10⁻⁶) and Saturday (−0.6374, OR ≈ 0.53, p = 7.1 × 10⁻⁸) are all significantly lower. The ordering is monotone, and the odds of clicking on a Saturday are roughly half those on a Monday.

**Week group (reference = Apr 6-11):** Clicking falls over the study. Apr 20-25 is indistinguishable from the reference (−0.0232, OR ≈ 0.98, p = 0.794), while May 4-9 (−0.1903, OR ≈ 0.83, p = 0.0451) and May 18-23 (−0.3435, OR ≈ 0.71, p = 0.0011) are significantly lower. This runs opposite to the completion model, where May 18-23 showed the highest completion: fewer riders opened the survey late in the study, but those who did were more likely to finish.

**Weekly ride count (−0.0076):** Each additional ride in the preceding seven days lowers the log-odds of clicking by 0.008 (OR ≈ 0.99), but the effect is not significant (p = 0.565). Frequency of service use shows no reliable association with the decision to open the survey, in contrast to the completion model where it was positive and significant.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# ---------------------------------------------------------------
# step 5 -  model 2: completion model, fit on CLICKERS ONLY
# F: binary indicator of completing the well-being question (not the survey) 
#   P(F = 1 | C = 1, I, R, U).  Covariates must be observed for
#   completers AND non-completers, so cannot be consequence
#   of completing -- so no progress, no duration, no trip_purpose.
# ---------------------------------------------------------------
clickers <- filter(riders, click == 1L)

m_main <- glm(complete ~  incentive + weekly_ride_count + hour + day + week_group ,
              family = binomial, data = clickers) #0-109, 1-1149
#m_int1 <- update(m_main, . ~ . + incentive:weekly_ride_count)
#m_int2 <- update(m_main, . ~ . + incentive:day)
#BIC(m_main, m_int1, m_int2)            # keep the lowest
fit_complete <- m_main                
summary(fit_complete)

#include the probability in riders dataframe
riders$prob_complete <- NA_real_
riders$prob_complete[riders$click == 1L] <-   # prob_complete has 8827 rows with NA
  predict(fit_complete, newdata = clickers, type = "response")
```

This is similar to the logistic model we just fitted.

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
# step 5:  joint probability & ipw weight.
#   P(F = 1, C = 1 | ) = P(F = 1 | C = 1, ) x P(C = 1 | )
#   A completer who was unlikely to reach this point receives a
#   large weight, standing in for similar riders who never did.

riders <- riders |>
  mutate(prob_click_and_complete = prob_click * prob_complete,
         ipw = 1 / prob_click_and_complete)

summary(riders$ipw[riders$complete == 1L])   # check for extreme weights
```

### 2.1 Theory of the inverse probability weight

**Notation.** Let $i = 1, \dots, N$ index the the rider-days on which a pop-up could have been seen ($N = 10{,}085$, contributed by 2,060 riders). For each opportunity define

- $R_i$: rider characteristics (day of week, hour of day),
- $U_i$: microtransit utilization (rides in the past seven days),
- $I_i$: the three-level incentive,
- $C_i \in \{0,1\}$: indicator of clicking on the pop-up,
- $F_i \in \{0,1\}$: indicator of completing the well-being question,
- $Y_i$: the well-being score.

Write $\mathbf{X}_i = (R_i, U_i)$ for the covariates observed at every opportunity, whether or not the rider clicked. Rider demographics $D$ are not recorded in either data set and so cannot enter the models.

**The selection mechanism.** The outcome $Y_i$ is observed only when a rider clears two hurdles in sequence: they must click the pop-up, and having clicked, they must answer the well-being question. Define the observation indicator

$$S_i = C_i \cdot F_i, \qquad Y_i \text{ observed} \iff S_i = 1.$$

The probability of being observed is therefore

$$\pi_i = P(S_i = 1 \mid \mathbf{X}_i, I_i) = P(C_i = 1, F_i = 1 \mid \mathbf{X}_i, I_i).$$

**Why the product of two probabilities:** The joint probability is not modelled directly. Applying the probability chain rule splits it into two pieces that can each be estimated from the data:

$$P(C = 1, F = 1 \mid \mathbf{X}, I)
= \underbrace{P(F = 1 \mid C = 1, \mathbf{X}, I)}_{\text{completion model}}
\times
\underbrace{P(C = 1 \mid \mathbf{X}, I)}_{\text{click model}}.$$

The second factor simplifies. Riders cannot see the incentive until after they click into the survey, so the incentive carries no information about the decision to click:

$$P(C = 1 \mid \mathbf{X}, I) = P(C = 1 \mid \mathbf{X}).$$

This identity is what makes the click layer estimable at all, the incentive is unobserved for riders who never clicked, and without the identity the first factor could not be fitted. Writing $\pi_i^{C}$ and $\pi_i^{F}$ for the two factors,

$$\pi_i = \pi_i^{C} \cdot \pi_i^{F},
\qquad
\pi_i^{C} = P(C_i = 1 \mid \mathbf{X}_i),
\qquad
\pi_i^{F} = P(F_i = 1 \mid C_i = 1, \mathbf{X}_i, I_i).$$

The two fitted models: Model 1 is fitted on all $N$ opportunities, and excludes the incentive by the identity above:

$$\operatorname{logit} \pi_i^{C}
= \alpha_0 + \alpha_1 U_i + \alpha_2 \text{hour}_i
+ \boldsymbol{\alpha}_3^\top \text{day}_i
+ \boldsymbol{\alpha}_4^\top \text{week}_i .$$

Model 2 is fitted on the clickers only, since $F$ is defined only once $C = 1$:

$$\operatorname{logit} \pi_i^{F}
= \gamma_0 + \boldsymbol{\gamma}_1^\top I_i + \gamma_2 U_i
+ \gamma_3 \text{hour}_i
+ \boldsymbol{\gamma}_4^\top \text{day}_i
+ \boldsymbol{\gamma}_5^\top \text{week}_i .$$

Every covariate in Model 2 must be observed for completers *and* non-completers, and must not be a consequence of completing. This is why $\textit{progress}$, $\textit{duration}$ and $\textit{trip\_purpose}$ are excluded: they are recorded only as a by-product of moving through the survey.

**The weight.** The inverse probability weight is the reciprocal of the estimated joint probability,

$$w_i = \frac{1}{\hat\pi_i}
= \frac{1}{\hat\pi_i^{C} \cdot \hat\pi_i^{F}} .$$

**Why inverting the probability corrects the bias:** The weight is the number of riders in the population that one observed rider stands in for. Following the Horvitz–Thompson argument, for any function $g(\cdot)$ of the outcome,

$$E\!\left[ \sum_{i=1}^{N} \frac{S_i}{\pi_i}\, g(Y_i) \right]
= \sum_{i=1}^{N} \frac{E[S_i \mid \mathbf{X}_i, I_i]}{\pi_i}\, g(Y_i)
= \sum_{i=1}^{N} \frac{\pi_i}{\pi_i}\, g(Y_i)
= \sum_{i=1}^{N} g(Y_i),$$

since $E[S_i \mid \mathbf{X}_i, I_i] = \pi_i$. The weighted sum over the observed riders is therefore an unbiased estimate of the sum over all riders, which is precisely the quantity of interest. Estimation proceeds by maximising the weighted log-likelihood over the observed cases,

$$\ell_w(\boldsymbol{\theta})
= \sum_{i \,:\, S_i = 1} \frac{1}{\hat\pi_i} \, \log L_i(\boldsymbol{\theta}),$$

which is what the `weights` argument supplies to the ordinal model in Step 6.

**Interpretation of the magnitude.** In this study $1{,}258$ of the $10{,}085$ opportunities produced a click, so $\hat P(C = 1) \approx 0.125$, and $1{,}149$ of those $1{,}258$ clickers answered the well-being question, so $\hat P(F = 1 \mid C = 1) \approx 0.913$. The joint probability is therefore

$$\hat\pi \approx 0.125 \times 0.913 \approx 0.114 \approx \frac{1}{9},
\qquad w = \frac{1}{\hat\pi} \approx 9 .$$

Roughly one rider in nine clears both hurdles, so each rider in the analysis represents about nine riders in Wilson. The weights are not constant: a rider whose covariate pattern makes clicking and completing unlikely receives a *larger* weight, because comparatively few riders like them survived both hurdles, and those few must therefore carry more of the load. Saturday illustrates this. Saturday riders click at roughly half the odds of Monday riders and complete slightly less often, so they are under-represented among the observed cases relative to their share of all opportunities; up-weighting them restores the composition of the full rider population.

**Assumptions.** Two conditions are required. First, positivity: every opportunity must have $\pi_i > 0$, so that no group of riders is structurally excluded from the sample. Second, and more restrictively, selection must be missing at random given the observed covariates,

$$S_i \perp\!\!\!\perp Y_i \mid \mathbf{X}_i, I_i .$$

That is, among riders sharing the same day, hour, recent ride count, week and incentive, whether a rider clicked and completed carries no further information about their well-being. If instead the decision to click or complete depends on current well-being itself, the mechanism is not missing at random and the weights cannot correct it, since they are constructed only from observed covariates. This limitation cannot be tested here, because the well-being score is missing precisely for those who did not complete. The absence of rider demographics $D$ further weakens the plausibility of the assumption, as age and gender are known predictors of survey non-response.

Weighting: each person you observe has to stand in for everyone like them that you didn't observe.

$$N \cdot \frac{1}{10} = n , \text{where } p = \frac{1}{10} \\
\text{30 individuals out of the sample were observed to have cars} \\
\text{ How many individuals in the population have cars ?} \\
\frac{30}{n} \cdot N = 30 \cdot (\frac{N}{n}) = 30 \cdot 10 = 300 \\
\text{Each samples person represents } \frac{1}{(1/p)} = 10 \text{ towns people.}$$

Why did we use the product of two probabilities?

A rider lands in the analysis only if they clear two hurdles:

- click on the POP-up

- complete the well-being question

$$P(C=1, F=1|cov) = 0.108 \approx \frac{1}{9} \\
\text{This implies that 1 in 9 riders make it to this sample.} \\
p = \frac{1}{9}, \textit{ipw} = \frac{1}{p} = 9 \
\text{Each sampled person represents 9 riders in Wilson} $$

So now n={clicking and completing the well-being question}

```{r, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}

# ---------------------------------------------------------------
# model 3: ipw-adjusted well-being model.
#   Fit on riders who clicked and completed.
#   day must stay in the model: each weekday asked about a different aspect of well-being, so dropping it would treat six distinct measures as if they were one.
# ---------------------------------------------------------------
analysis <- riders |>
  filter(complete == 1L, !is.na(wb)) |>
  mutate(wb_ord = factor(wb, levels = 1:7, ordered = TRUE))

fit_wb_ipw <- MASS::polr(wb_ord ~ weekly_ride_count + day,
                         data = analysis, weights = analysis$ipw, Hess = TRUE)

# Unweighted fit on the same completers, for comparison
# the gap between the two columns shows how much the weighting moves the estimates.
fit_wb_cc  <- MASS::polr(wb_ord ~ weekly_ride_count + day,
                         data = analysis, Hess = TRUE)

cbind(complete_case = coef(fit_wb_cc), ipw = coef(fit_wb_ipw))

# ---------------------------------------------------------------
# Step 7:  Repeated-measures version.
#   The simulation has one row per rider; ride_data has 1,441
#   surveys from 736 riders (up to 16 each), so a rider random
#   intercept is needed.  polr cannot fit one; clmm can.
# ---------------------------------------------------------------
fit_wb_clmm <- clm(wb_ord ~ weekly_ride_count + day, #+ (1 | unique_id)
                    data = analysis, weights = ipw, link = "logit")
summary(fit_wb_clmm)
```

```{r}
fit_wb_clmm <- clm(wb_ord ~ weekly_ride_count, #+ (1 | unique_id)
                   data = analysis, weights = ipw, link = "logit",
                   subset = day == "Monday")
summary(fit_wb_clmm)
```
